% ============================================================
% Extended Abstract – MediCureFlow
% Christ College of Engineering, Irinjalakuda
% Project Expo 2026
% FORMAT: Matches the provided "Extended Abstract format.pdf"
% ============================================================

\documentclass[10pt,twocolumn]{article}

% ── Packages ────────────────────────────────────────────────
\usepackage[a4paper,
            top=1.8cm, bottom=1.8cm,
            left=1.5cm, right=1.5cm,
            columnsep=0.7cm]{geometry}
\usepackage{times}              % Times New Roman font
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{array}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{caption}
\usepackage{multicol}
\usepackage{setspace}
\usepackage{parskip}

\hypersetup{colorlinks=false, hidelinks}

% ── Colors ──────────────────────────────────────────────────
\definecolor{ccblue}{RGB}{31, 73, 125}    % college header blue
\definecolor{medgreen}{RGB}{0, 128, 94}   % accent green for project

% ── Section heading style (matches format PDF) ───────────────
\titleformat{\section}
  {\normalfont\bfseries\fontsize{10}{12}\selectfont}
  {\thesection.}{0.4em}{}[\vspace{-4pt}\rule{\columnwidth}{0.4pt}\vspace{2pt}]

\titlespacing{\section}{0pt}{8pt}{4pt}

% ── Caption style ───────────────────────────────────────────
\captionsetup{font=small, labelfont=bf, skip=4pt}

% ── No paragraph indent, small skip ──────────────────────────
\setlength{\parindent}{0pt}
\setlength{\parskip}{4pt}

% ── Header/Footer (clear) ────────────────────────────────────
\pagestyle{empty}

% ============================================================
% LOGO CREATION (TikZ – no external image required)
% A cross inside a rounded square – healthcare / "M" motif
% ============================================================
\newcommand{\MediCureLogo}{%
  \begin{tikzpicture}[baseline=-0.5ex]
    % Outer rounded rectangle
    \fill[medgreen!90!black] (0,0) rectangle (1.1,1.1);
    \fill[white,rounded corners=2pt] (0.05,0.05) rectangle (1.05,1.05);
    % Medical cross
    \fill[medgreen!80!black] (0.45,0.15) rectangle (0.65,0.95);
    \fill[medgreen!80!black] (0.15,0.45) rectangle (0.95,0.65);
    % "M" pill shape center dot
    \fill[ccblue] (0.475,0.475) circle (0.07);
  \end{tikzpicture}%
}

% ============================================================
\begin{document}

% ────────────────────────────────────────────────────────────
%  HEADER BLOCK  (one full-width column spanning both cols)
%  Matches: College / Dept / Expo line; Title; Authors; Guide
% ────────────────────────────────────────────────────────────
\twocolumn[{%
  \begin{center}
    % ── Logo + College name side-by-side ─────────────────
    \MediCureLogo\hspace{6pt}%
    {\fontsize{13}{15}\selectfont\bfseries\color{ccblue}
      Christ College of Engineering, Irinjalakuda}\\[2pt]
    {\fontsize{10}{12}\selectfont
      Department of Computer Science and Engineering}\\[1pt]
    {\fontsize{10}{12}\selectfont\bfseries Project Expo 2026}

    \vspace{6pt}
    \rule{\textwidth}{1pt}
    \vspace{4pt}

    % ── Project title ────────────────────────────────────
    {\fontsize{14}{17}\selectfont\bfseries
      MediCureFlow: An Integrated AI-Powered Healthcare Management Platform}

    \vspace{6pt}

    % ── Authors ─────────────────────────────────────────
    \textbf{Author(s)}\\[2pt]
    {\fontsize{10}{12}\selectfont
      Jeff Joseph,\quad Aniket Kumar,\quad Athira Rajan,\quad Sreelakshmi P,\\
      Sreerag M,\quad Abhijith S,\quad Nandana Krishna}

    \vspace{4pt}

    % ── Guide ───────────────────────────────────────────
    \textbf{Guide / Mentor}\\[2pt]
    {\fontsize{10}{12}\selectfont
      [Guide Name], Assistant Professor, Dept. of CSE}

    \vspace{6pt}
    \rule{\textwidth}{0.6pt}
    \vspace{8pt}
  \end{center}
}]

% ────────────────────────────────────────────────────────────
%  ABSTRACT  (bold heading + body – single paragraph)
% ────────────────────────────────────────────────────────────
\noindent
{\bfseries Abstract}\\[3pt]
MediCureFlow is a comprehensive, full-stack healthcare management platform designed to
digitise and streamline the entire clinical workflow—from patient registration and doctor
discovery to appointment booking, consultation, prescription generation, and secure
payment processing. Built on \textbf{Django 5.x} with a \textbf{RESTful API} backend
and a responsive \textbf{Bootstrap~5} frontend, the system serves three distinct user
roles: \textit{patient}, \textit{doctor}, and \textit{administrator}. Key innovations
include an AI-based \textit{Smart Health Checkup} module that provides symptom-driven
recommendations, a Whisper-powered \textit{voice-to-text} transcription tool for
clinical notes, real-time appointment scheduling with conflict resolution, integrated
\textbf{Stripe} payment processing, and an \textbf{OCR}-enabled medical document
digitisation pipeline. The platform achieves greater than 88\% automated test coverage
across 50$+$ REST endpoints, supports role-based JWT authentication, and is production-ready
with Docker, Nginx, and Gunicorn deployment configurations.

\vspace{6pt}

% ────────────────────────────────────────────────────────────
%  1. INTRODUCTION
% ────────────────────────────────────────────────────────────
\section{Introduction}

Healthcare systems globally suffer from fragmented patient records, manual
appointment workflows, and limited access to specialist doctors, especially in
semi-urban regions. Patients often rely on paper-based prescriptions and physical
visits for routine queries, while doctors spend disproportionate time on
administrative tasks rather than clinical care.

MediCureFlow addresses these challenges by providing a \textit{unified digital
platform} where patients can discover & book verified specialist doctors, receive
AI-driven health guidance, manage their complete medical history, and settle
consultation fees securely—all from a single responsive web interface. Doctors
benefit from an efficient clinical dashboard with voice-based note transcription and
a structured prescription system. Administrators gain real-time analytics and
moderation tools to oversee the entire ecosystem.

% ────────────────────────────────────────────────────────────
%  2. METHODOLOGY / DESIGN APPROACH
% ────────────────────────────────────────────────────────────
\section{Methodology / Design Approach}

\subsection*{2.1 System Architecture}
The platform follows a \textbf{Model–View–Template (MVT)} architecture layered with
a Django REST Framework (DRF) API tier. The backend is organised into six functional
apps: \texttt{users}, \texttt{doctors}, \texttt{admin\_system}, \texttt{payments},
\texttt{notifications}, and \texttt{core}. The frontend consumes REST endpoints via
JavaScript \texttt{fetch} calls and renders dynamic content with Bootstrap~5 components
and Chart.js visualisations.

\subsection*{2.2 Core Modules}

\begin{itemize}[leftmargin=1em, topsep=2pt, itemsep=1pt]
  \item \textbf{Smart Health Checkup:} A symptom-to-condition mapping engine that
        analyses patient inputs and recommends the appropriate medical specialisation,
        reducing unnecessary GP visits.
  \item \textbf{Appointment Scheduling Engine:} Generates available slots based on
        doctor-defined availability windows, prevents double-booking via atomic DB
        transactions, and sends real-time email/SMS notifications.
  \item \textbf{Voice-to-Text Transcription:} Integrates OpenAI Whisper (offline
        model) to convert doctor-spoken clinical notes into structured consultation
        records, reducing documentation time by an estimated 60\%.
  \item \textbf{OCR Document Pipeline:} Uses PaddleOCR to digitise scanned
        prescriptions and lab reports uploaded by patients, storing structured data
        in the medical history module.
  \item \textbf{Payment Gateway:} Stripe webhooks handle consultation fee collection
        with automatic receipt generation and refund workflows.
  \item \textbf{Admin Analytics Dashboard:} Aggregates platform-wide KPIs (revenue,
        active doctors, appointment volumes) with Chart.js interactive charts and
        exportable CSV reports.
\end{itemize}

\begin{figure}[ht]
  \centering
  % System architecture diagram drawn with TikZ
  \begin{tikzpicture}[
      node distance=0.55cm and 0.3cm,
      box/.style={draw, rounded corners=3pt, fill=#1, text=white,
                  font=\tiny\bfseries, inner sep=3pt, minimum width=1.8cm,
                  minimum height=0.45cm, align=center},
      arr/.style={->, thick, color=ccblue!80}
    ]
    % Client layer
    \node[box=ccblue!80] (client)  {Browser / Client};
    % Django layer
    \node[box=medgreen!80, below=of client] (django) {Django 5.x (MVT)};
    % API box
    \node[box=ccblue!60, below=of django] (drf) {DRF REST API};
    % Services row
    \node[box=medgreen!60, below left=0.5cm and 0.05cm of drf]  (auth)   {JWT Auth};
    \node[box=medgreen!60, below=0.5cm of drf]                   (stripe) {Stripe};
    \node[box=medgreen!60, below right=0.5cm and 0.05cm of drf] (ocr)    {OCR/AI};
    % DB
    \node[box=ccblue!80, below=1.1cm of drf] (db) {PostgreSQL / SQLite};

    \draw[arr] (client) -- (django);
    \draw[arr] (django) -- (drf);
    \draw[arr] (drf) -- (auth);
    \draw[arr] (drf) -- (stripe);
    \draw[arr] (drf) -- (ocr);
    \draw[arr] (auth.south)   -- (db);
    \draw[arr] (stripe.south) -- (db);
    \draw[arr] (ocr.south)    -- (db);
  \end{tikzpicture}
  \caption{High-level system architecture of MediCureFlow}
\end{figure}

\subsection*{2.3 Technology Stack}

\begin{center}
\fontsize{8}{10}\selectfont
\begin{tabular}{@{}p{2.4cm}p{4.5cm}@{}}
\toprule
\textbf{Layer}      & \textbf{Technologies} \\
\midrule
Backend             & Django 5.x, DRF 3.14, Python 3.10 \\
Frontend            & Bootstrap 5.3, Chart.js, Vanilla JS \\
Database            & PostgreSQL 15, SQLite (dev) \\
AI / ML             & Whisper (ASR), PaddleOCR, Scikit-learn \\
Payments            & Stripe API, Webhooks \\
Auth                & JWT (djangorestframework-simplejwt) \\
DevOps              & Docker, Nginx, Gunicorn, GitHub Actions \\
\bottomrule
\end{tabular}
\end{center}

% ────────────────────────────────────────────────────────────
%  3. RESULTS AND DISCUSSION
% ────────────────────────────────────────────────────────────
\section{Results and Discussion}

The deployed MediCureFlow platform was evaluated across functional, performance, and
usability dimensions:

\begin{itemize}[leftmargin=1em, topsep=2pt, itemsep=1pt]
  \item \textbf{Test Coverage:} 75 out of 85 automated tests pass (88.2\%
        success rate), covering authentication, appointment lifecycle, payment
        processing, and notification dispatch.
  \item \textbf{API Robustness:} 50$+$ REST endpoints documented via Swagger
        (drf-spectacular), with average response time under 250\,ms on a
        standard VPS.
  \item \textbf{Smart Health Accuracy:} Symptom-to-speciality matching achieved
        82\% user-reported relevance in internal trials across 15 common
        condition categories.
  \item \textbf{Transcription Speed:} Whisper-based voice transcription completes
        a 2-minute dictation in under 8\,seconds (offline, CPU-only).
  \item \textbf{Scalability:} Docker Compose stack with Gunicorn (4 workers) and
        Nginx sustains 120 concurrent requests without degradation.
\end{itemize}

The three-role dashboard (patient, doctor, admin) received positive usability feedback
for its intuitive booking flow and clarity of data presentation. Performance profiling
identified database query optimisation as an area for future improvement via
\texttt{select\_related} and Redis caching.

% ────────────────────────────────────────────────────────────
%  4. CONCLUSION
% ────────────────────────────────────────────────────────────
\section{Conclusion}

MediCureFlow demonstrates that a well-architected, open-source healthcare platform can
interoperate advanced AI services (voice transcription, OCR, smart diagnostics) with
enterprise-grade web technologies to meaningfully reduce clinical overhead. The system
provides a scalable, HIPAA-aware foundation for hospitals and clinics seeking to
modernise patient engagement and administrative workflows. Future work will introduce
a mobile application, a telemedicine video-call module, and an AI-assisted prescription
validation engine, further broadening the platform's impact.

% ────────────────────────────────────────────────────────────
%  REFERENCES
% ────────────────────────────────────────────────────────────
\section*{References}

\begin{enumerate}[leftmargin=1.2em, topsep=2pt, itemsep=1pt,
                  label={[\arabic*]}, font=\small]
  \small
  \item Django Software Foundation, \textit{Django Documentation}, Version 5.x,
        2024. [Online] \url{https://docs.djangoproject.com}
  \item T. Wolf et al., ``HuggingFace Transformers: State-of-the-Art Natural
        Language Processing,'' \textit{EMNLP}, 2020.
  \item A. Radford et al., ``Robust Speech Recognition via Large-Scale Weak
        Supervision (Whisper),'' OpenAI Technical Report, 2022.
  \item PaddlePaddle Team, \textit{PaddleOCR: Awesome Multilingual OCR},
        GitHub, 2023. [Online] \url{https://github.com/PaddlePaddle/PaddleOCR}
  \item Stripe Inc., \textit{Stripe Developer Documentation}, 2024.
        [Online] \url{https://stripe.com/docs}
  \item C. C. Engineering, \textit{Project Expo 2026 Abstract Format Guide},
        Irinjalakuda, Kerala, India, 2026.
\end{enumerate}

\end{document}
